Temporal Developer Concept Witchery: Latent Dirichlet Allocation Applied to Revisions

### 2008-10-25 lda-paper.tex
2008-10-24 21:59:21 abez

\documentclass{report}
\usepackage{graphicx,amsmath,comment,fullpage}
\newcommand{\igH}[1]{\includegraphics[height=.9\textheight]{#1}}
\newcommand{\igW}[1]{\includegraphics[width=.9\textwidth]{#1}}
\newcommand{\igWh}[1]{\includegraphics[width=.7\textwidth]{#1}}
\newcommand{\igWhalf}[1]{\includegraphics[width=.45\textwidth]{#1}}
\newcommand{\lda}{Latent Dirichlet Allocation}

\author{Abram Hindle\\
ahindle@cs.uwaterloo.ca\\
Software Architecture Group (SWAG)\\
Supervised by\\
Micheal Godfrey and Ric Holt
}
\title{Lab 0006: Latent Dirichlet Allocation Applied to Revisions}

@@

\subsection{Intro to LDA}

XXX explain LDA \cite{944937}

\subsection{LDA Work}

Lukins, Kraft and Etzkorn~\cite{lukins2008} use LDA to help query for
finding related documents during bug localization tasks. They used LDA
to build a topic model and then query against it using sample
documents, the query would return relevant documentation.

LSI is related to LDA and has been used to identify topics in software
artifacts~\cite{1421013,1374321,10.1109/ICPC.2007.13,10.1109/ICPC.2006.17}.

Semantic Clustering \cite{1698774,1566153}

Linstead et al has proposed an author source code model using
LDA.\cite{10.1109/MSR.2007.20,NIPS2007637,1321709}

Grant et al. have used an alternative technique to LSI and LDA, they
used Independant Component Analysis to seperate topic signals from
software code.

@@

\section{First Impression}

% 2007-10-12

Looking at 30 day non-overlapping windows of revisions for MySQL 3.23
(20 topics, top 10 words per topic) we found there were common words
across topic clusters such as diffs, annotate and history. There were
notable transitional topics such as in the first window the word
bitkeeper appears (probably when they adopted bitkeeper) yet in the
following windows there were no mentions of bitkeeper. ``RENAME'' also
only appeared once in the first window.

Often words are shared across topics, see table \ref{tab:portability}
for tracking topics about portability.

It is apparent that we have to compare these topics some how by
itemset or similarity.

\begin{table}
\centering
\begin{tabular}{|ccc|l|}
\hline
2000 &  Jul &  31 &    chmod \\
2000 &  Sep &  29 &    fixes benchmark logging windows \\
2000 &  Nov &  28 &    typo fix insert\_multi\_value \\
2001 &  Jan &  27 &    FIxes Innobase Cleanups auto-union \\
2001 &  Mar &  28 &    2 topics bugfix, logging , TEMPORARY,  \\
\hline
2001 &  Jul &  26 &    Update Allow TABLES LOCK [a] \\

2001 &  Aug &  25 &    Tables row version [a] \\
\hline
2001 &  Sep &  24 &    Update checksum merge \\
2001 &  Oct &  24 &    fixed fix \\
2001 &  Dec &  23 &    HPUX SCO fix \\
\hline
2002 &  Feb &  21 &    net buffer length  max\_allowed\_packet [b] \\
2002 &  Mar &  23 &    small buf fix [b]  \\
\hline
2002 &  May &  22 &    [popular] fix SCO OSF1 table\_name \\
2002 &  Nov &  18 &    HPUX11 compiler HP \\
\hline
2003 &  Feb &  16 &    Linux errno  [c] \\
2003 &  Mar &  18 &    alarm bookmark bug [c] \\
\hline
2003 &  Sep &  14 &    Auto logging merge windows distribution fix 64-bit 4.0 Cleanup \\
\hline
\end{tabular}
\caption{Tracking topics associated with the word portability, note some continuous blocks}
\label{tab:portability}
\end{table}

@@

\bibliographystyle{alpha}
\bibliography{paper1}

>>>>

\documentclass[times, 10pt,twocolumn]{article} 
\usepackage{times}

\usepackage{graphicx,amsmath,comment,fullpage}
\newcommand{\igH}[1]{\includegraphics[height=.9\textheight]{#1}}
\newcommand{\igW}[1]{\includegraphics[width=.9\textwidth]{#1}}
\newcommand{\igWh}[1]{\includegraphics[width=.7\textwidth]{#1}}
\newcommand{\igWhalf}[1]{\includegraphics[width=.45\textwidth]{#1}}
\newcommand{\lda}{Latent Dirichlet Allocation}

\author{Abram Hindle and Michael Godfrey and Ric Holt \\
\{ahindle,migod,holt\}@cs.uwaterloo.ca\\
Software Architecture Group (SWAG)\\
University of Waterloo\\
}
\title{Temporal Developer Concept Witchery: Latent Dirichlet Allocation Applied to Revisions}

@@

%\subsection{Intro to LDA}

%XXX explain LDA \cite{944937}

Latent Dirichlet Alloation (LDA)~\cite{944937} is a competitor to
latent semantic indexing (LSI), probabilistic latent semantic indexing
(pLSI) and semantic clustering. It is used for document modelling,
document clustering and collaborative filtering. What LDA attempts to
do is model a documents as mixture model of topics.  The common use of
LDA in software engineering
literature~\cite{lukins2008,10.1109/MSR.2007.20,NIPS2007637,1321709}
is to extract topic clusters from documents such as methods, bug
reports, source code and files.

Our blackbox view of LDA and LSI is that they extract topic clusters
from the documents that are independant from one another. LDA doesn't
name these clusters but the clusters consist of words whose
relationship is often obvious to someone familar with the
corpus. Technically we could swap LDA for LSI. Our data is posed as
documents with word distributions (word counts per documents) and LDA
extracts distinct topics (clusters of words) from the documents.

\subsection{LDA, LSI and Semantic Clustering}

Linstead et al has proposed an author source code model using
LDA~\cite{10.1109/MSR.2007.20,NIPS2007637,1321709}. They used an
author-topic (AT) model~\cite{1036902} to infer the association or
expertise of a particular developer with a particular code topic
extracted from the source control system.

Lukins, Kraft and Etzkorn~\cite{lukins2008} use LDA to help query for
finding related documents during bug localization tasks. They used LDA
to build a topic model and then query against it using sample
documents, the query would the proportion of topics that were
associated with the query and thus the related documents.

LSI is related to LDA and has been used to identify topics in software
artifacts for forumal concept analysis and concept
location~\cite{1421013,1374321,10.1109/ICPC.2007.13,10.1109/ICPC.2006.17}.
This is when concepts are automatically extracted from source code or
documents are related to part of the source code.  Concept location is
how high level concepts relate to low level entities like source code,
for instance when fixing a bug how do the concepts in the bug report
map back to the source code?  Semantic Clustering has also been used
for similar reason~\cite{1698774,1566153}.

Grant et al.~\cite{scottcordy} have used an alternative technique to
LSI and LDA, they used Independant Component Analysis to seperate
topic signals from software code.

@@

\section{First Impressions}

In our first exploratory pass we wanted to see

Looking at 30 day non-overlapping windows of revisions for MySQL 3.23
(20 topics, top 10 words per topic) we found there were common words
across topic clusters such as diffs, annotate and history. There were
notable transitional topics such as in the first window the word
``bitkeeper'' appears (probably when they adopted Bitkeeper for their
source control system) yet in the following windows there were no
mentions of Bitkeeper. ``RENAME'' also only appeared once in the first
window and never again.

A subsampling of the notable topic words is displayed in Table
\ref{tab:portability}. We tracked one topic cluster which seemed
related to portability.

%Often words are shared across topics, see table \ref{tab:portability}
%for tracking topics about portability.

After viewing these topic is was obvious that we would have to combine
try to track the evolution of topics over the windows or rank topics
by similarity.

%It is apparent that we have to compare these topics some how by
%itemset or similarity.

\begin{table*}
\centering
\begin{tabular}{|ccc|l|}
\hline
2000 &  Jul &  31 &    chmod \\
2000 &  Sep &  29 &    fixes benchmark logging windows \\
2000 &  Nov &  28 &    typo fix insert\_multi\_value \\
2001 &  Jan &  27 &    fixes Innobase Cleanups auto-union \\
2001 &  Mar &  28 &    2 topics bugfix, logging , TEMPORARY,  \\
\hline
2001 &  Jul &  26 &    update Allow TABLES LOCK [a] \\

2001 &  Aug &  25 &    tables row version [a] \\
\hline
2001 &  Sep &  24 &    update checksum merge \\
2001 &  Oct &  24 &    fixed fix \\
2001 &  Dec &  23 &    HPUX SCO fix \\
\hline
2002 &  Feb &  21 &    net buffer length  max\_allowed\_packet [b] \\
2002 &  Mar &  23 &    small buf fix [b]  \\
\hline
2002 &  May &  22 &    [popular] fix SCO OSF1 table\_name \\
2002 &  Nov &  18 &    HPUX11 compiler HP \\
\hline
2003 &  Feb &  16 &    Linux errno  [c] \\
2003 &  Mar &  18 &    alarm bookmark bug [c] \\
\hline
2003 &  Sep &  14 &    Auto logging merge windows distribution fix 64-bit 4.0 Cleanup \\
\hline
\end{tabular}
\caption{Tracking topics associated with the word portability, note some continuous blocks}
\label{tab:portability}
\end{table*}

@@

\bibliographystyle{alpha}
\bibliography{lda-paper}
